To run with the hare and hunt with the hounds: Optimal recommendation system with competing sellers. \\ Master Thesis

### 2020-04-19 Thesis.tex
Add files via upload
Thesis

\documentclass[a4paper]{article}
\usepackage[12pt]{extsizes} % 
\usepackage{setspace}
\doublespacing
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
%\usepackage{harvard}
\usepackage[round]{natbib}
\usepackage{pdfpages}
\usepackage{tikz}
\usepackage{makecell}
\usepackage{amsthm}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage{breakcites}
\interfootnotelinepenalty=10000
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
%\newenvironment{abstract}[0]{\small\rm
%	\begin{center}ABSTRACT
%		\\ \vspace{8pt}
%		\begin{minipage}{5.2in}\smalllineskip
%			\hspace{1pc}}{\end{minipage}\end{center}\vspace{-1pt}}
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=25mm, top=20mm, right=25mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{definition}[theorem]{Definition}
%\let\cite = \citeasnoun
\title{To run with the hare and hunt with the hounds: Optimal recommendation system with competing sellers. \\ Master Thesis}
\date{}
\author{Danil Fedchenko}%\thanks{I am very grateful to Sergei Izmalkov for his help during the work on this project}}
\begin{document} % 
	\maketitle
	\begin{abstract}
		Many e-commerce platforms that connect buyers and sellers employ recommender systems to help customers find products and services. These systems may have several potential effects on strategies of competing sellers: on one hand, more people become aware of products, so sellers may sell more goods and benefit from the presence of the recommender system. On the other hand, at the same time, more people become aware of the competitor seller. The latter effect may intensify the price competition and may hurt sellers but, of course, benefit customers. However, many platforms charge firms for sales and as a result have the stakes in firms profits. Thus, by designing the recommender policy the platform should balance the pro-competitive care about customers and anti-competitive incentives to keep high prices. The main goal of this work is to study this trade-off in details. I am solving for an optimal recommender policy and investigate factors that may affect it.
	\end{abstract}
\newpage
\tableofcontents
\newpage
	\section{Introduction}

		Suppose you came to New-York and would like to find a good restaurant to go out tonight. Or you are going to spend your upcoming vacation on the coast of Spain and you need to find a hotel. What would you do? What will be the first thing you interact with?  Before the World Wide Web, people in such cases relied upon guide-books and brochures distributed by travel agents. Those guides usually quickly became out-of-date and were criticized for containing misleading information.

@@

In the rows of examples, the special attention should be paid to online retailers such as \textit{Amazon}, \textit{eBay}, \textit{AliExpress} and many others. These platforms stand at the forefront of online Economics, generating huge amounts of money and completely transforming the shopping experience for people all other the world. These platforms play roles of intermediaries between buyers and sellers, helping the buyers to find the goods they need. And this process clearly includes making recommendations which seller to display you when you are searching for something.

@@

For example, consider a retailer (such as \textit{Amazon} or \textit{eBay}), these retailers charge sellers commissions for each sale. These commissions usually take the form of a percentage commission. That means that the retailer obtains a higher profit if the price of the deal is higher. So, one of the questions that could potentially arise then is: should, in this case, the recommender system which Amazon employs out of two sellers who sell (almost) substitutable goods recommend the one who has a higher price? Or suppose that \textit{eBay} is thinking on the commission rate it charges (now it is about 8\%), should the commission be set at the highest possible level that guarantees that sellers do not refrain from using the platform? What if, at the same time, the quality of goods may differ and the sellers may have an option to spend some resources on disclosing this quality (e.g. to make the detailed description or take high-resolution photos). Clearly the higher is the commission the less willing are the sellers to spend money on quality disclosure: cost does not outweigh benefits. But if sellers do not disclose the information this may adversely affect the customers' satisfaction since this may create a mismatch between buyers and sellers: people end up with the good which does not fit them well. Thus, thinking about commission rate the platform should take this aspect into account. Additionally, suppose that the recommendation system is not precise i.e. it sometimes makes mistakes on figuring out the customer's preferences. How does this inaccuracy affect the behavior (price setting) of competing sellers and how does it interact with their incentives to disclose the quality of their goods? What should be the optimal commissions and recommender policy in this case? Although in this work I am touching upon all questions mentioned above, the main objective of this paper is to understand what should be the optimal recommendation strategy and commission rate the platform charges the competitive sellers, with the particular emphasis on the interaction with sellers' incentives to disclose private information about their goods.

	To address these questions I build a theoretical one-period model (please see the detailed description of the model in the corresponding section). I consider the unit mass of heterogeneous customers whose preferences are summarized by the point on this unit line. At the extremes of the unit interval, there are two sellers. Customers demand one good and derive utility upon consuming the good but bear ``transportation'' cost by purchasing. This cost, of course, has little to do with any transportation but rather represents the outside option, opportunity cost of the customer: the utility the customer should refrain from in order to consume the good. I suppose that the qualities of goods offered by sellers can be different and that these qualities are \textit{ex-ante} privately known only to the corresponding sellers who can costly disclose this information before setting prices. The main player in this system is the recommender system (the platform) who beforehand receives (possibly noisy) signal regarding each customer's location and then may recommend any of the two sellers to this customer. This feature of my model has clear real-life interpretation: location of the customer represents his or her tastes, the signal which reveals the location can be interpreted as the result of some fancy machine-learning algorithm the majority of platform nowadays use in order to assess, to estimate the true preferences, tastes of the customer. The platform also sets a commission rate (percentage) it charges the sellers for each sale. However, I explicitly assume that the platform cares not only about the profit (commissions) but also about customers i.e. consumer surplus directly enters the platform's objective function. I suppose that this assumption is not very far away from reality: clearly, if the platform cares only about the profit and this adversely affects customers' satisfaction and happiness from using the service, in the long-run people simply will refuse using it or switch to the competitive platforms. The latter suggests that for the long-run sustainability the platform should pay some attention to the consumer surplus.

In the work, I find the optimal recommender policy and investigate the effect of the presence of the recommender system on equilibrium behavior of sellers. I demonstrate that the recommender system which has stakes in the sellers' profits softens competition. The latter effect leads to higher prices. The amount of information disclosed in equilibrium is determined in the interplay between two effects: on one hand, higher commission directly decreases profits and, hence, incentives of sellers to disclose the information; on the other hand, the higher commission increases platform's stakes in profits and softens competition which results in higher prices and higher profits. The total value of these two effects is determined by the degree of consumer-orientedness of the platform: if it is high enough, increase in prices does not outweigh direct losses, so the total effect is negative; in the opposite case total effect is positive and an increase in commission increases the amount of information disclosed.

In the extension of the baseline model, I allow for the not perfectly precise recommender system. Namely, I assume that the signal, the platform receives reveals the true location of the customer only with some probability $\beta$, and with probability $1 - \beta$ the signal follows a uniform distribution. This has several consequences: first of all, now the platform adjusts its recommender strategy incorporating this inaccuracy. This adjustment manifests itself in the way how the platform treats and processes the signal regarding the location, now the platform takes into consideration that the signal could potentially be wrong. The latter effectively increases the share of customers in the middle, increasing incentives for sellers to undercut the competitor and to get access to this increased demand. Thus, the direct effect of the platform's inaccuracy intensifies the competition between sellers. This decreases the sellers' profits and as a result, decreases sellers' incentives to reveal the quality. At the same time, the recommender system's inaccuracy increases the attention the platform pays to the quality of the good when it makes recommendations. The latter effect increases the sellers' returns on quality disclosure making the latter more profitable. Eventually, the total effect also depends on parameters: if the cost of disclosure is high enough, the first effect dominates and improving the recommender system precision decreases the amount of information the sellers disclose, for the low cost of disclosure the second effect prevails and increasing the platform's precision increases the amount of information disclosed.

@@

Three papers mentioned above although address the question on optimal recommender policy, do not take into account the strategic nature of recommended alternatives.
\cite{chen2016advertising} consider the informative role of recommender system on strategies of competing sellers and the equilibrium outcomes. They consider two sellers located at the extremes of Hotelling-like (\cite{harold1929stability}) linear city who are competing for the unit mass of customers, located in between. There is also a recommender system that receives noisy information about each customer's location and can recommend one of the sellers to the customers. Although in this paper, the recommendation strategy is exogenous, the authors consider how the presence of the recommender system affects competition between sellers. In particular, they are looking at prices and advertising expenditure. The main finding of this paper is that the presence of the recommender system may fundamentally change the nature of competition. In particular, the authors demonstrate that with the recommender system prices and advertising expenditure switch from being strategic substitutes to strategic complements. This results in the fact that with recommender system sellers prefer to advertise less, relying more on advertising provided by the recommendation system (the authors called this effect \textit{advertising effect}). At the same time, prices with recommender system are lower (the authors called this effect \textit{competition effect}) which is the result of tougher competition. Interestingly, the total effect of recommender system on sellers profits becomes a result of an interplay between \textit{competition effect} and \textit{advertising effect}: for some values of parameters the sellers become worse-off with the presence of the recommender system, \textit{competition effect} dominates, while for other values the sellers benefit from ``free'' advertising provided by the platform.

@@

\cite{levin2009quality} add a vertical component, focusing on the sellers' decisions regarding the quality disclosure, although, in the framework without the recommender system. In the model which I consider, I firstly endogenise the recommendation strategy that the platform implements, and secondly, I am looking at the interplay between the recommendation policy and sellers' incentives to disclose the information about the goods' quality.	
	\section{Model}
The model I consider is similar to that considered by \cite{levin2009quality}, the difference is that I introduce a new player -- a platform which employs a recommender system (I will use words ``platform'' and ``recommender system'' interchangeably).

There is a unit mass of heterogeneous customers. Preferences of customer are characterized by the location of the customer on the unity interval. Two competing sellers (firms) which I will refer to $A$ and $B$, who locate at the extremes of the unit interval, are selling substitutable goods which they produce at equal marginal cost normalized to 0. Goods of each seller may have a different quality which means that the customers with the same ``horizontal preference'' for the two sellers may derive different utility upon consuming goods of different sellers. Sellers are competing in prices which they set simultaneously. In addition, the sellers may costly disclose the information about the quality of their good prior to price setting. The recommender system observes prices and sellers' decisions regarding the quality disclosure and makes a recommendation to customers which seller to choose. In the baseline model, I assume that the recommender system knows the exact location of the customer, lately, this assumption is relaxed, and I consider the scenario when the recommender system receives a noisy signal regarding the customers' locations. Finally, the platform charges sellers a commission for sales according to a commission rate $\alpha$ which the platform can choose.

@@

	\subsection{Customers}
Customer who locates at $x \in [0, 1]$ derives a net utility $$V + q_A - p_A - x$$ if  he buys the good at a price $p_A$ from the seller $A$ who locates at a point $0$, and net utility $$V + q_B - p_B - (1-x)$$ if  he buys the good at a price $p_B$ from the seller $B$ who locates at a point $1$. $V$ is assumed to be higher than $\frac{3}{2}$ to ensure that the market is fully covered in equilibrium, $q_A, q_B \in [0, 1]$ are the qualities of the goods. The outside option for the customers is zero, so they follow the recommendation if the expected net utility from buying the good from the recommended seller is higher than 0.

Customers do not know about the sellers unless the recommender system recommends them.

@@

The platform commits to the commission rate $\alpha$ and the recommender policy. The recommender policy specifies functions $r_A(x, p_A, p_B, d_A, d_B, q_A, q_B)$ and $r_B(x, p_A, p_B, d_A, d_B, q_A, q_B)$ which represents probabilities that the customer with location $x$ will be recommended the seller $A$ and $B$ respectively. The objective function of the platform is given by \begin{align}\label{pl_profit}
\pi_P = \alpha(p_A D_A + p_B D_B) + \gamma CS
\end{align}

where $CS$ is a consumer surplus, $\gamma$ represents the platform's degree of consumer-orientedness. I assume that the maximum rate of commission the platform can charge the sellers is $\bar{\alpha} \in (0, 1)$, and $\gamma > \bar{\alpha}$.

@@

\section{Results}
Firstly, I consider the case with observable qualities and without the recommender system. I need this comparison in order to demonstrate how and via which mechanisms the presence of the recommender system affects the equilibrium behavior of sellers.

@@

	\begin{figure}[H]
		\centering
		\includegraphics[width=0.6\textwidth]{ImpactRS}
		\caption{Impact of RS $(p_A < p_B)$}\label{fig1}
	\end{figure}
	As we can see, there exists a share of customers $\Delta = \frac{\alpha}{2 \gamma} (p_B - p_A)$ such that they strictly prefer the firm $A$ to the firm $B$ while are recommended the firm $B$. This happens since from the platform's standpoint the losses in consumer surplus are lower than gains driven by high-price sales. Note, also, that $\frac{\partial \Delta}{\partial \alpha} > 0$, $\frac{\partial \Delta}{\partial \gamma} < 0$ and $\lim_{\gamma \to \infty} \Delta = 0$. That is, the impact of the platform eliminates completely if $\gamma$ becomes very high, and the more the platform is profit-oriented (the higher is $\alpha$) the higher is $\Delta$.
	\subsection{Qualities are not observable}
Here I consider the case when qualities $q_A, q_B$ are not observable to the platform unless the sellers decide to disclose it. In this subsection, I derive the optimal recommender system of the platform. Following \cite{levin2009quality} I consider symmetric PBE in which the sellers disclose the quality if and only if the quality falls above some threshold $q^*$ and do not disclose otherwise, and if the seller does not disclose, the beliefs of the platform regarding the quality do not depend on prices.

@@

Formally, I assume that before making recommendation to a particular customer whose true location $x$ is equal to $x_0 \in [0, 1]$ the platform receives a signal $s$ such that $$\mathbb{P}(s=x_0|x=x_0) = \beta \in [0, 1]$$ and with probability $1-\beta$, $s \sim U(0, 1)$. I will refer to $\beta$ as the recommender system precision. In this case, as it is shown by \cite{li2018recommender} the conditional expectation regarding the true location is equal to \begin{align}\label{signal}
\mathbb{E}[x|s=x_0] = \beta x_0 + \frac{1-\beta}{2}
\end{align}
The result above is quite intuitive: the higher is precision the more weight the platform assigns to the signal, while when the precision is low the platform more frequently infers that the signal is not correct and prescribes to the customer an average location $\frac{1}{2}$.

>>>>

\documentclass[a4paper]{article}
\usepackage[12pt]{extsizes} % 
\usepackage{setspace}
\doublespacing
\usepackage[utf8]{inputenc}
\usepackage{pgf,tikz,pgfplots}
\usetikzlibrary{patterns}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
%\usepackage{harvard}
\usepackage[round]{natbib}
\usepackage{pdfpages}
\usepackage{tikz}
\usepackage{makecell}
\usepackage{amsthm}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage{breakcites}
\interfootnotelinepenalty=10000
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
%\newenvironment{abstract}[0]{\small\rm
%	\begin{center}ABSTRACT
%		\\ \vspace{8pt}
%		\begin{minipage}{5.2in}\smalllineskip
%			\hspace{1pc}}{\end{minipage}\end{center}\vspace{-1pt}}
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=25mm, top=20mm, right=25mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{definition}[theorem]{Definition}
%\let\cite = \citeasnoun
\title{To run with the hare and hunt with the hounds: Optimal recommendation system with competing sellers. \\ Master Thesis}
\date{}
\author{Danil Fedchenko\thanks{I am very grateful to: Sergei Izmalkov for constant help and support during the work on this project; Marta Troya Martinez for many valuable comments and suggestions. \newline Any remaining errors are mine.}}
\begin{document} 
	\maketitle
	\begin{abstract}
		Many e-commerce platforms that connect buyers and sellers employ recommender systems to help customers find products and services. These systems may have several potential effects on strategies of competing sellers: on one hand, more people become aware of products, so sellers may sell more goods and benefit from the presence of the recommender system. On the other hand, at the same time, more people become aware of the competitor seller. The latter effect may intensify the price competition and may hurt sellers but, of course, benefit customers. However, many platforms charge firms for sales and as a result have the stakes in firms profits. Thus, by designing the recommender policy the platform should balance the pro-competitive care about customers and anti-competitive incentives to keep high prices. The main goal of this work is to study this trade-off in details. I am solving for an optimal recommender policy and investigate factors that may affect it.
	\end{abstract}
\newpage
\tableofcontents
\newpage
	\section{Introduction}

		Suppose you came to New-York and would like to find a good restaurant to go out tonight. Or you are going to spend your upcoming vacation on the coast of Spain and you need to find a hotel. Or you would like to buy something using your favorite online retailer. What would you do? What will be the first thing you interact with?  Before the World Wide Web, to find hotels or restaurants people relied upon guide-books and brochures distributed by travel agents. Those guides usually quickly became out-of-date and were criticized for containing misleading information.

@@

In the rows of examples, the special attention should be paid to online retailers such as \textit{Amazon}, \textit{eBay}, \textit{AliExpress} and many others. These platforms stand at the forefront of online Economics, generating huge amounts of money\footnote{e.g. \textit{Amazon}'s gross merchandise value (total value of goods sold on the platform) in 2019 fiscal year account for \$335 billion.} and completely transforming the shopping experience for people all other the world. These platforms play roles of intermediaries between buyers and sellers, helping the buyers to find the goods they need. And this process clearly includes making recommendations which seller to display you when you are searching for something.

@@

For example, consider a retailer (such as \textit{Amazon} or \textit{eBay}), these retailers charge sellers commissions for each sale. These commissions usually take the form of a percentage commission. That means that the retailer obtains a higher profit if the price of the deal is higher. So, one of the questions that could potentially arise then is: should, in this case, the recommender system which \textit{Amazon} employs out of two sellers who sell (almost) substitutable goods recommend the one who has a higher price? Or suppose that \textit{eBay} is thinking on the commission rate it charges, %(now it is about 8\%),
 should the commission be set at the highest possible level that guarantees that sellers do not refrain from using the platform? What if, at the same time, the quality of goods may differ and the sellers may have an option to spend some resources on disclosing this quality (e.g. to make the detailed description or take high-resolution photos). Clearly the higher is the commission the less willing are the sellers to spend money on quality disclosure: cost does not outweigh benefits. But if sellers do not disclose the information this may adversely affect the customers' satisfaction since this may create a mismatch between buyers and sellers: people end up with the good which does not fit them well. Thus, thinking about commission rate the platform should take this aspect into account. Additionally, suppose that the recommendation system is not precise i.e. it sometimes makes mistakes on figuring out the customer's preferences. How does this inaccuracy affect the behavior (price setting) of competing sellers and how does it interact with their incentives to disclose the quality of their goods? What should be the optimal commissions and recommender policy in this case? Although in this work I am touching upon all questions mentioned above, the main objective of this paper is to understand what should be the optimal recommendation strategy and commission rate the platform charges the competitive sellers, with the particular emphasis on the interaction with sellers' incentives to disclose private information about their goods.

	To address these questions I build a theoretical one-period model. I consider a unit mass of heterogeneous customers whose preferences are summarized by the point on this unit line. At the extremes of the unit interval, there are two sellers. Customers demand one good and derive utility upon consuming the good but bear ``transportation'' cost by purchasing. This cost although has nothing to do with transportation can be interpreted in several possible ways. For example, it can be think of as cost of the outside option, opportunity cost of the customer: the utility the customer should refrain from in order to consume the good; or as the measure of quality of fit: how well the product fits the particular customer. The latter allows to treat the location of the customer as the customer's taste regarding the two goods offered by sellers: closer the customer is to the seller $A$ -- better the good of seller $A$ fits this customer, other things being equal. Moreover, I suppose that the qualities of goods offered by sellers can be different. In other words, all customers regardless of their location get the same additional value upon consuming the good with higher quality. I assume that these qualities are \textit{ex-ante} privately known only to the corresponding sellers who can costly disclose this information before setting prices. As I mentioned earlier, the cost of disclosure may include time and monetary cost of making a detailed description of the good, or taking many high-resolution photos of the item. The main player in this interaction is the recommender system (the platform) who beforehand receives (possibly noisy) signal regarding each customer's location and then may recommend any of the two sellers to this customer. This feature of my model has clear real-life interpretation: the signal which reveals the location can be interpreted as the result of some fancy machine-learning algorithm the majority of platform nowadays use in order to assess, to estimate the true preferences, tastes of the customer. The platform also sets a commission rate (percentage) it charges the sellers for each sale. However, I explicitly assume that the platform cares not only about the profit (commissions) but also about customers i.e. consumer surplus directly enters the platform's objective function. I suppose that this assumption is not very far away from reality: clearly, if the platform cares only about the profit and this adversely affects customers' satisfaction and happiness from using the service, in the long-run people simply will refuse using it or switch to the competitive platforms. The latter suggests that for the long-run sustainability the platform should pay some attention to the consumer surplus.

	Also, I argue that the model I consider, as a whole, fits well the research questions I am addressing. The recommender systems usually help when customers are horizontally differentiated: there can be 10 restaurants of top quality, but you may love only one of them simply because it fits your tastes better (e.g. you just like Indian food). In this case, a good recommender system should recommend you precisely this restaurant out of ten. If, on the contrary, everybody agreed that restaurant $A$ is in all respects worse than the restaurant $B$, nobody would need the recommender system.

In the work, I find the optimal recommendation policy and investigate the effect of the presence of the recommender system on equilibrium behavior of sellers. I demonstrate that the recommender system which has stakes in the sellers' profits softens competition. The latter effect leads to higher prices. The amount of information about the quality of goods, disclosed in equilibrium is determined in the interplay between two effects: on one hand, higher commission directly decreases profits and, hence, incentives of sellers to disclose the information; on the other hand, the higher commission increases platform's stakes in profits and softens competition which results in higher prices and higher profits. The total value of these two effects is determined by the degree of consumer-orientedness of the platform: if it is high enough, increase in prices does not outweigh direct losses, so the total effect on the amount of disclosed information is negative; in the opposite case total effect is positive and an increase in commission increases the amount of information disclosed.

In the extension of the baseline model, I allow for the not perfectly precise recommender system. Namely, I assume that the signal, the platform receives reveals the true location of the customer only with some probability $\beta$, and with probability $1 - \beta$ the signal follows a uniform distribution. This has several consequences: first of all, now the platform adjusts its recommender strategy incorporating this inaccuracy. This adjustment manifests itself in the way how the platform treats and processes the signal regarding the location, now the platform takes into consideration that the signal could potentially be wrong. In this case, the best prediction of the customer's location the platform could made is simply average location. The latter effectively increases the share of customers in the middle of the unity interval, increasing incentives for sellers to undercut the competitor and to get access to this increased demand. Thus, the direct effect of the platform's inaccuracy intensifies the competition between sellers. This decreases the sellers' profits and as a result, decreases sellers' incentives to reveal the quality. At the same time, the recommender system's inaccuracy increases the attention the platform pays to the quality of the good when it makes recommendations: since the platform assumes that horizontal component (location) may be measured imprecisely, making a recommendation it begins to put more weights on vertical component (quality) which, if disclosed, is precise. The latter effect increases the sellers' returns on quality disclosure making the latter more profitable. Eventually, the total effect also depends on parameters: if the cost of disclosure is high enough, the first effect dominates and improving the recommender system precision decreases the amount of information the sellers disclose, for the low cost of disclosure the second effect prevails and increasing the platform's precision increases the amount of information disclosed.

@@

Three papers mentioned above although address the question on optimal recommender policy, do not take into account the strategic nature of recommended alternatives.
\cite{chen2016advertising} consider the informative role of recommender system on strategies of competing sellers and the equilibrium outcomes. They consider two sellers located at the extremes of Hotelling-like (\cite{harold1929stability}) linear city who are competing for the unit mass of customers, located in between. There is also a recommender system that receives noisy information about each customer's location and can recommend one of the sellers to the customers. Although in this paper, the recommendation strategy is exogenously given and fixed, the authors consider how the presence of the recommender system affects competition between sellers. In particular, they are looking at prices and advertising expenditure. The main finding of this paper is that the presence of the recommender system may fundamentally change the nature of competition. In particular, the authors demonstrate that with the recommender system prices and advertising expenditure switch from being strategic substitutes to strategic complements. This results in the fact that with recommender system sellers prefer to advertise less, relying more on advertising provided by the recommendation system (the authors called this effect \textit{advertising effect}). At the same time, prices with recommender system are lower (the authors called this effect \textit{competition effect}) which is the result of tougher competition. Interestingly, the total effect of recommender system on sellers profits becomes a result of an interplay between \textit{competition effect} and \textit{advertising effect}: for some values of parameters the sellers become worse-off with the presence of the recommender system, \textit{competition effect} dominates, while for other values the sellers benefit from ``free'' advertising provided by the platform.

@@

\cite{levin2009quality} add a vertical component to the standard Hotelling model, focusing on the sellers' decisions regarding the quality disclosure, although, in the framework without the recommender system. In the model which I consider, I firstly endogenise the recommendation strategy that the platform implements, and secondly, I am looking at the interplay between the recommendation policy and sellers' incentives to disclose the information about the goods' quality.	
	\section{Model}
The model I consider is similar to that considered by \cite{levin2009quality}, the difference is that I introduce a new player -- a platform which employs a recommender system (I will use words ``platform'' and ``recommender system'' interchangeably).

There is a unit mass of heterogeneous customers. Preferences of customer are characterized by the location of the customer on the unity interval. Two competing sellers (firms) which I will refer to $A$ and $B$, who locate at the extremes of the unit interval, are selling substitutable goods which they produce at equal marginal cost normalized to 0. Goods of each seller may have a different quality which means that the customers with the same ``horizontal preference'' for the two sellers may derive different utility upon consuming goods of different sellers. Sellers are competing in prices which they set simultaneously. In addition, the sellers may costly disclose the information about the quality of their good to the platform prior to price setting. The recommender system observes prices and sellers' decisions regarding the quality disclosure and makes a recommendation to customers which seller to choose. In the baseline model, I assume that the recommender system knows the exact location of the customer, lately, this assumption is relaxed, and I consider the scenario when the recommender system receives a noisy signal regarding the customers' locations. Finally, the platform charges sellers a commission for sales according to a commission rate $\alpha$ which the platform can choose.

@@

	\subsection{Customers}
Customer who locates at $x \in [0, 1]$ derives a net utility $$V + q_A - p_A - x$$ if  he buys the good at a price $p_A$ from the seller $A$ who locates at a point $0$, and net utility $$V + q_B - p_B - (1-x)$$ if  he buys the good at a price $p_B$ from the seller $B$ who locates at a point $1$. $V$ is assumed to be higher than $\frac{3}{2}$ to ensure that the market is fully covered in equilibrium, $q_A, q_B \in (0, 1)$ are the qualities of the goods. These qualities are \textit{ex-ante} known only to corresponding sellers, and are $i.i.d.\  U(0, 1)$. The outside option for the customers is zero, so they follow the recommendation if the expected net utility from buying the good from the recommended seller is higher than 0.

Customers do not know about the sellers (hence cannot sell) unless the recommender system recommends them.

@@

The platform commits to the commission rate $\alpha$ and the recommender policy. The recommender policy specifies functions $r_A(x, p_A, p_B, d_A, d_B, \mathbb{E}q_A, \mathbb{E}q_B)$ and $r_B(x, p_A, p_B, d_A, d_B, \mathbb{E} q_A, \mathbb{E} q_B)$ which represents probabilities that the customer with location $x$ will be recommended the seller $A$ and $B$ respectively. The objective function of the platform is given by \begin{align}\label{pl_profit}
\pi_P = \mathbb{E}\left[\alpha(p_A D_A + p_B D_B) + \gamma CS\right]
\end{align}

where expectation is taken over unknown qualities $q_A, q_B$; $CS$ is a consumer surplus, $\gamma$ represents the platform's degree of consumer-orientedness. I assume that the maximum rate of commission the platform can charge the sellers is $\bar{\alpha} \in (0, 1)$, and $\gamma > \bar{\alpha}$. Usually in real life there are several competing platforms, so the platform cannot increase the commission up to 100\% since in this case sellers will just switch to the competitive online retail platform. So, $\bar{\alpha}$ can be think of as a maximum commission rate the platform can charge that still guarantees that sellers are willing to sell their goods through the platform.

@@

\section{Results}
Firstly, I consider the case with observable to buyers qualities and without the recommender system. I need this comparison in order to demonstrate how and via which mechanisms the presence of the recommender system affects the equilibrium behavior of sellers.

@@

\begin{figure}[H]
	\centering
	\begin{tikzpicture}[scale=1.5]
	\draw [fill = black] (0, 0) rectangle (6.5, 0.03);
	\draw (6.5, -0.4) node[color=red]{$B$};
	\draw (0, 0.34) node[]{$0$};
	\draw (6.5, 0.34) node[]{$1$};
	\draw (-0.05, -0.4) node[color=blue]{$A$};
	\draw (3.25, -0.47) node[]{$\frac{1}{2}$};
	\draw (3.65, -0.4) node[]{$\hat{x}$};
	\draw (4.46, -0.4) node[]{ $x_0$};
	\draw [fill = black] (0, -0.1) rectangle (0.02, 0.12);
	\draw [fill = black] (3.24, -0.1) rectangle (3.26, 0.12);
	\draw [fill = black] (6.48, -0.1) rectangle (6.5, 0.12);
	\draw [fill = black] (3.64, -0.1) rectangle (3.66, 0.12);
	\draw [fill = black] (4.45, -0.1) rectangle (4.47, 0.12);
	\draw[pattern=north west lines, pattern color=blue] (0,0) rectangle (3.64,0.2);
	\draw[pattern=north west lines, pattern color=red] (3.66,0) rectangle (6.48,0.2);
	\draw[pattern=north west lines, pattern color=blue] (0,0) rectangle (4.45,-0.2);
	\draw[pattern=north west lines, pattern color=red] (4.47,0) rectangle (6.48,-0.2);
	\draw [dashed] (3.65,0) -- (3.65,0.7);
	\draw [dashed] (4.46,0) -- (4.46,0.7);
	\draw [<->] (3.65,0.7) -- (4.46,0.7);
	\draw (4.05, 0.9) node[]{$\Delta$};
	\end{tikzpicture}
	\caption{Impact of RS $(p_A < p_B)$}\label{fig1}
\end{figure}
	As we can see, there exists a share of customers $\Delta = \frac{\alpha}{2 \gamma} (p_B - p_A)$ such that they strictly prefer the firm $A$ to the firm $B$ while are recommended the firm $B$. This happens since from the platform's standpoint the losses in consumer surplus are lower than gains driven by high-price sales. Note, also, that $\frac{\partial \Delta}{\partial \alpha} > 0$, $\frac{\partial \Delta}{\partial \gamma} < 0$ and $\lim_{\gamma \to \infty} \Delta = 0$. That is, the impact of the platform eliminates completely if $\gamma$ becomes very high, and the more the platform is profit-oriented (the higher is $\alpha$) the higher is $\Delta$.
	\subsection{Qualities are not observable}
Here I consider the case when qualities $q_A, q_B$ are not observable to the platform unless the sellers decide to disclose it. In this subsection, I derive the optimal recommender system of the platform. Following \cite{levin2009quality} I consider symmetric PBE in which the sellers disclose the quality if and only if the quality falls above some threshold $q^*$ and do not disclose otherwise, and if the seller does not disclose, the beliefs of the platform regarding the quality do not depend on prices.

@@

Formally, I assume that before making recommendation to a particular customer whose true location $x$ is equal to $x_0 \in [0, 1]$ the platform receives a signal $s$ such that $$\mathbb{P}(s=x_0|x=x_0) = \beta \in [0, 1]$$ and with probability $1-\beta$, $s \sim U(0, 1)$. I will refer to $\beta$ as the recommender system precision. In this case, it can be shown (see  \cite{li2018recommender} for example) that the conditional expectation regarding the true location is equal to \begin{align}\label{signal}
\mathbb{E}[x|s=x_0] = \beta x_0 + \frac{1-\beta}{2}
\end{align}
The result above is quite intuitive: the higher is precision the more weight the platform assigns to the signal, while when the precision is low the platform more frequently infers that the signal is not correct and prescribes to the customer an average location $\frac{1}{2}$.
